---
output:
  word_document: default
  html_document: default
  pdf_document: default
---
```{r}
# Part A: 
covid_dataset <- read.csv("UScovid.csv")

# filter covid_dataset
covid_dataset <- covid_dataset[covid_dataset$date == "2021-06-03" & covid_dataset$county != "Unknown", ]

# removes "date" and "fips"
covid_dataset <- covid_dataset[, !(names(covid_dataset) %in% c("date", "fips"))]

# alphebectial order
covid_dataset <- covid_dataset[order(covid_dataset$state, covid_dataset$county), ]


head(covid_dataset)
```


```{r}
# Part B: Death Rate column calulated from latest
covid_dataset$death.rate <- round((covid_dataset$deaths / covid_dataset$cases) * 100, 2)

head(covid_dataset)

```



```{r}
# Part C: Top 10 countries with largest number of cases
top_cases <- covid_dataset[order(covid_dataset$cases, decreasing = TRUE), ][1:10, c("state", "county", "cases", "deaths", "death.rate")]
print(top_cases)
```
```{r}
# Part D: Top 10 counties with the largest number of deaths
top_deaths <- covid_dataset[order(covid_dataset$deaths, decreasing = TRUE), ][1:10, c("state", "county", "cases", "deaths", "death.rate")]
print(top_deaths)
```
```{r}
#Part E : Top 10 counties with the highest case fatality rates

top_fatality <- covid_dataset[order(covid_dataset$death.rate, decreasing = TRUE), ][1:10, c("state", "county", "cases", "deaths", "death.rate")]

print(top_fatality)
```

```{r}
#Part F: Top 10 counties with the highest case fatality rates and at least 100,000 cases

top_100k_cases <- covid_dataset[covid_dataset$cases >= 100000, ]

top_fatality_100k <- top_100k_cases[order(top_100k_cases$death.rate, decreasing = TRUE), ][1:10, c("state", "county", "cases", "deaths", "death.rate")]

print(top_fatality_100k)


```
```{r}
# Part G: Albemarle, Virginia

print(covid_dataset[covid_dataset$county == "Albemarle" & covid_dataset$state == "Virginia", c("county", "state", "cases", "deaths", "death.rate")])


```
```{r}
# Part G: Charlottesville city, Virgnia

print(covid_dataset[covid_dataset$county == "Charlottesville city" & covid_dataset$state == "Virginia", c("county", "state", "cases", "deaths", "death.rate")])
```

```{r}
#Part A: State Level Anaylsis
state_leveldata <- aggregate(cbind(cases, deaths) ~ state, data = covid_dataset, sum)


state_leveldata <- state_leveldata[order(state_leveldata$state), ]


head(state_leveldata)
```
```{r}
# Part B: Calculate Case Fatality Rate at the state level
state_leveldata$state.rate <- round((state_leveldata$deaths / state_leveldata$cases) * 100, 2)

# Display the first 6 rows of the state-level data with the new death rate column
head(state_leveldata)

```

```{r}
#Part (c) - Case Fatality Rate for Virginia
virginia_rate <- state_leveldata[state_leveldata$state == "Virginia", "state.rate"]
print(virginia_rate)


```
```{r}
# 3. Part D: Case Fatality Rate for Puerto Rico
puerto_rico_rate <- state_leveldata[state_leveldata$state == "Puerto Rico", "state.rate"]
print(puerto_rico_rate)
```

```{r}
# Part E: The highest top 10 fatality rate
top_states_fatality <- state_leveldata[order(state_leveldata$state.rate, decreasing = TRUE), ][1:10, ]
print(top_states_fatality)

```
```{r}
# Part F: States with the 10 lowest fatality 
bottom_states_fatality <- state_leveldata[order(state_leveldata$state.rate, decreasing = FALSE), ][1:10, ]
print(bottom_states_fatality)
```
```{r}
#CSV export
write.csv(state_leveldata, "stateCovid.csv", row.names = FALSE)
```
```
```{r}


```

1.What is the case fatality rate in Virginia? the case fertility rate was 1.66

2.What is the case fatality rate in Puerto Rico? The case fertility rate for Puerto Rico was 0. This was not in the dataset.

3. Which states have the 10 highest case fatality rate? 
Massachusetts	
New Jersey
New York
Connecticut	
District of Columbia
Mississippi
Pennsylvania
Louisiana	
New Mexico
Maryland

4.Which states have the 10 lowest case fatality rate?
Alaska	
Utah
Virgin Islands	
Vermont	
Nebraska	
Idaho	
Northern Mariana Islands	
Wisconsin	
Wyoming
Colorado

